---
output:
  md_document:
    variant: markdown_github
---

#BBGDM
[![Travis-CI Build Status](https://travis-ci.org/skiptoniam/bbgdm.svg?branch=master)](https://travis-ci.org/skiptoniam/bbgdm)

BBGDM is a R package for running Generalized Dissimilarity Models with Bayesian Bootstrap for parameter estimation. To install package run the following command in your R terminal
```{r,eval=FALSE}
install.packages(c('devtools'))
devtools::install_github('skiptoniam/bbgdm')
```

### Load the required libaries, we need vegan for the dune dataset.
```{r,error=FALSE,message=FALSE,warning=FALSE}
library(bbgdm)
library(vegan)
```

### Run bbgdm on the famous dune meadow data
The dune meadow vegetation data, dune, has cover class values of 30 species on 20 sites.
Make the abundance data presence/absence.
```{r}
data(dune)
data(dune.env)
dune_pa <- ifelse(dune>0,1,0)
```

### Fit a bbgdm
Now we have a species by sites matrix of simulated data and a set data for a one dimensional gradient.
```{r,message=FALSE,warning=FALSE,error=FALSE}
form <- ~1+A1
fm1 <- bbgdm(form,dune_pa, dune.env,family="binomial",link='logit',
             dism_metric="number_non_shared",spline_type = 'ispline',
             nboot=100, geo=FALSE,optim.meth='nlmnib')
```

### Plot diagnostics
```{r,fig.width = 4, fig.height = 4}
resids <- diagnostics(fm1)
par(mfrow=c(2,2))
plot(resids)
```


### Plot response curves
```{r,fig.width = 4, fig.height = 4}
response <- as.response(fm1)
par(mfrow=c(1,1))
plot(response)
```

### Run 'Wald-like' test on parameters
```{r,results='asis'}
library(xtable)
wt <- bbgdm.wald.test(fm1)
tab <- xtable(wt)
print(tab, type = "html")
```
