---
title: "Bayesian Bootstrap Generalised Dissimilarity Modelling Vignette"
author: "Skipton Woolley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Here we present a basic vignette that runs through taking a species by sites matrix and performs a Bayesian Bootstrap Generalised Dissimilarity Model. Our vignette will run through the following steps:

- Simulate a dataset of species and  covariates for fitting.
- Fit a bbgdm model to this these data.
- Run diagnostic, inference and plotting functions.

## Generate some simulated data
Generate species with responses to covariates and to mix it up (skew the distributions).

```{r loadlibraries, echo = FALSE, warning=FALSE, message=FALSE}
library(bbgdm)
```

```{r simulate data, warning=FALSE, message=FALSE}
mean1 <- c(1.5,1.5) #for intercepts and slopes # a hypothesis on the response of each parameter...
mean2 <- c(-1.5,-1.5)
means <- rbind(mean1,mean2)
variances <- matrix(c(rep(.56,dim(means)[2]),rep(.56,dim(means)[2])),dim(means)[1],dim(means)[2],byrow=T)
variances <- t(variances)
covariances <- c(-.5,-.5) #apply(variances,1,function(x)1/(sqrt(mean(x))*sqrt(mean(x))))
mix.prop <- 0.5
nSp <- 200   #lots of species (to see pattern in distribution)
nSites <-50
dat <- data.frame(y=rep(1,nSites),x=seq(-2,2,length.out = nSites))

species_theta_generation_mixed_proportions <- function(means,variances,covariances,
                                                       nSp,dat,mix.prop,dist='negbin',
                                                       phi=NULL,plot=TRUE){
  nSigma <- dim(means)[1]
  dimSigma <- dim(means)[2]
  sigmas <- array(0L,c(dimSigma,dimSigma,nSigma))  
  species_parameters <- array(0L,c(nSp,dimSigma,nSigma))
  for(i in 1:nSigma){
    tmp <-sigmas[,,i] 
    tmp <- diag(variances[i,])
    tmp[row(tmp)!=col(tmp)] <- covariances[i]
    sigmas[,,i]<-tmp
    species_parameters[,,i] <- mvtnorm::rmvnorm(nSp, means[i,], sigmas[,,i])
  }
  grp <- rbinom(nSp, c(nSigma)-1, mix.prop)
  df <- data.frame(grp=grp,do.call(cbind,plyr::alply(species_parameters,3)))
  theta <- matrix(0L,nSp,dimSigma)
  for(i in 1:nSigma){
    row_tmp <- which(grp == c(i-1))
    theta[row_tmp,] <- as.matrix(subset(df, grp == c(i-1), select=c(c(i+1):c(i+nSigma))))
  }
  out <- matrix(0, dim(dat)[1], nSp)
  X <- as.matrix(dat)
  for (s in 1:nSp){
    if (dist == "bernoulli") {
      lgtp <- X %*% theta[s, ]
      p <- exp(lgtp)/(1 + exp(lgtp))
      out[, s] <- rbinom(dim(X)[1], 1, p)
    }
    if (dist == "negbin") {
      if(is.null(phi))phi<-rep(1,nSp)
      tmp <- rep(1e+05, dim(X)[1])
      while (max(tmp, na.rm = T) > 50000 | sum(tmp) < 2) {
        lpd_sp <- X %*% theta[s,]  
        p <- exp(lpd_sp)
        tmp<-rnbinom(dim(X)[1],mu=p,size=1/phi[s])
      }
      out[,s] <- tmp
    }
    if (dist == "poisson") {
      tmp <- rep(1e+05, dim(X)[1])
      while (max(tmp, na.rm = T) > 50000 | sum(tmp) < 2) {
        lpd_sp <- X %*% theta[s,]  
        p <- exp(lpd_sp)
        tmp<-rpois(dim(X)[1],lambda =p)
      }
      out[,s] <- tmp
    }
  }
  if(plot)for(i in 1:dim(theta)[2]-1)filled.contour(MASS::kde2d(theta[,1], theta[,i+1]), xlab="Intercepts", ylab=paste0("Slopes",i))
  return(list(sp_data=out, thetas=theta, mu=means,sigma=sigmas))
}

set.seed(42)
sim_data <- species_theta_generation_mixed_proportions(means,variances,covariances,nSp,dat,mix.prop,dist='bernoulli',plot=TRUE)
```

```{r fit model, results="hide", error=FALSE, warning=FALSE, message=FALSE}
form <- ~1+x
fm1_nlm <-bbgdm::gdm.bb(form,sim_data$sp_data,dat,nboot = 100,geo = FALSE,spline_type = "ispline",optim.meth = 'nlmnib', est.var = TRUE)
```

```{r, fig.show='hold' ,echo = FALSE, fig.height=6,fig.width=8}
plotResponse(fm1_nlm,plotdim = c(1,1))
```
