---
title: "Bayesian Bootstrap Generalised Dissimilarity Modelling Vignette"
author: "Skipton Woolley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run simple BBGDM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Here we present a basic vignette that runs through taking a species by sites matrix and performs a Bayesian Bootstrap Generalised Dissimilarity Model. Our vignette will run through the following steps:

- Fit a bbgdm model to a community dataset.
- Run diagnostic, inference and plotting functions.

### Install bbgdm from github, this requires devtools to be installed

```{r}
# install.packages(c('devtools','vegan'))
# devtools::install_github('skiptoniam/bbgdm')
```

### Load the required libaries, we need vegan for the dune dataset.
```{r loadlibraries, echo = TRUE, error=FALSE, warning=FALSE, message=FALSE}
library(bbgdm)
library(vegan)
```

### Run bbgdm on the famous dune meadow data
The dune meadow vegetation data, dune, has cover class values of 30 species on 20 sites.
Make the abundance data presence/absence.
```{r dune data, error=FALSE, warning=FALSE, message=FALSE}
data(dune)
data(dune.env)
dune_pa <- ifelse(dune>0,1,0)
```

### Fit a bbgdm
Now we have a species by sites matrix of simulated data and a set data for a one dimensional gradient.
```{r fit model, results="hide", error=FALSE, warning=FALSE, message=FALSE}
form <- ~1+A1
fm1 <- bbgdm(form,dune_pa, dune.env,family="binomial",link='logit',
             dism_metric="number_non_shared",spline_type = 'ispline',
             nboot=100, geo=FALSE,optim.meth='nlmnib')
```

### Plot diagnostics
```{r,fig.width = 6, fig.height = 6, fig.align='center'}
resids <- diagnostics(fm1)
par(mfrow=c(2,2))
plot(resids)
```

### Plot response curves
```{r,fig.width = 4, fig.height = 4, fig.align='center'}
response <- as.response(fm1)
par(mfrow=c(1,1))
plot(response)
```

### Run 'Wald-like' test on parameters
```{r,results='asis'}
library(xtable)
wt <- bbgdm.wald.test(fm1)
tab <- xtable(wt)
print(tab, type = "html")
```

### References
Jongman, R.H.G, ter Braak, C.J.F & van Tongeren, O.F.R. (1987). Data Analysis in Community and Landscape Ecology. Pudoc, Wageningen.
