---
title: "Bayesian Bootstrap Generalised Dissimilarity Modelling Vignette"
author: "Skipton Woolley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run simple BBGDM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


Here we present a basic vignette that runs through taking a species by sites matrix and performs a Bayesian Bootstrap Generalised Dissimilarity Model. Our vignette will run through the following steps:

- Fit a bbgdm model to a community dataset.
- Run diagnostic, inference and plotting functions.

### Install bbgdm from github, this requires devtools to be installed

```{r}
# install.packages(c('devtools','vegan'))
# devtools::install_github('skiptoniam/bbgdm')
```

### Load the required libaries, we need vegan for the dune dataset.
```{r loadlibraries, echo = TRUE, error=FALSE, warning=FALSE, message=FALSE}
library(bbgdm)
library(vegan)
```

### Run bbgdm on the famous dune meadow data
The dune meadow vegetation data, dune, has cover class values of 30 species on 20 sites.
Make the abundance data presence/absence.
```{r dune data, error=FALSE, warning=FALSE, message=FALSE}
data(dune)
data(dune.env)
dune_pa <- ifelse(dune>0,1,0)
```

### Fit a bbgdm
Now we have a species by sites matrix of simulated data and a set data for a one dimensional gradient.
```{r fit model, results="hide", error=FALSE, warning=FALSE, message=FALSE}
form <- ~1+A1
fm1 <- bbgdm(form,dune_pa, dune.env,family="binomial",dism_metric="number_shared",
             nboot=100, scale_covar=F,geo=F,optim.meth='nlmnib')
```

### Plot response curves
```{r, fig.show='hold' ,echo = TRUE, fig.height=6,fig.width=8}
plotResponse(fm1,plotdim = c(1,1))
```
### Plot diagnostics
```{r, fig.show='hold' ,echo = TRUE, fig.height=6,fig.width=8}
bbgdm.check(fm1)
```
### Run 'Wald-like' test on parameters
```{r}
bbgdm.wald.test(fm1)
```

### References
Jongman, R.H.G, ter Braak, C.J.F & van Tongeren, O.F.R. (1987). Data Analysis in Community and Landscape Ecology. Pudoc, Wageningen.
